---
title: "CGM"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data

read in CGM data and retreieve genes associated with GO term. I picked GO:0007601(visual perception) since it has 200 genes

functions 

```{r message=FALSE , results = 'hide'}

library(tidyverse)
library(caret)
library(PRROC)
library(mlbench)
set.seed(121212)
setwd('~/NIH/CGM/')

read_data <- function(filename,go_id){
    data<- read_tsv(filename) %>% dplyr::select(-ENSGene_Name) %>% .[!duplicated(.['Gene_Name']),] %>% as.data.frame()
    any(is.na(data))
    mart <- biomaRt::useMart(biomart = 'ENSEMBL_MART_ENSEMBL', dataset = 'hsapiens_gene_ensembl')
    gene_list <- biomaRt::getBM(attributes = 'external_gene_name',filters = 'go',values = go_id,mart = mart)%>%pull(external_gene_name)
    data$in_go_set <- ifelse(data$Gene_Name%in%gene_list,'in_set','not_in_set')
    data <- data %>% select(-Gene_Name)
}
run_analysis <- function(data, model, tr_ctrl){
    #proably doing this wrong
   idx=createDataPartition(data$in_go_set,p=.75)[[1]]
   data <- preProcess(data,method=c("center", "scale"))%>%predict(data)
   d_train=data[idx,]
   d_test=data[-idx,]
   set.seed(121212)
   fit <- train(in_go_set ~ . , data = d_train,
                method= model,
                trControl=tr_ctrl)
   prediction_table <- data.frame(in_go_set=d_test$in_go_set, predict(fit,d_test, type = 'prob'), 
                                  prediction=predict(fit,d_test))# yup im lazy
   index_1 <- prediction_table$in_go_set=='in_set'
   index_2 <- prediction_table$in_go_set=='not_in_set'
   
   cm <- confusionMatrix(prediction_table$prediction,d_test$in_go_set%>%as.factor)
   pr_plot=pr.curve(prediction_table$in_set[index_1],prediction_table$in_set[index_2],curve = T)
   return(list(prediction_table,cm,pr_plot))
}
compare_sampling <- function(data,model){
    df=data.frame()
    samp_types <- c('none','down','up','smote')
    for(samp in samp_types){
        tr_ctrl <- trainControl(method = 'cv', number = 5, classProbs = T,sampling = samp)
        if(samp == 'none') tr_ctrl <- trainControl(method = 'cv', number = 5, classProbs = T)
        res <- run_analysis(data = data,model = model, tr_ctrl = tr_ctrl)
        k= res[[3]]$curve %>% as.data.frame %>% rename(Precision=V1,Recall=V2) %>% mutate(samp=samp, auc=res[[3]]$auc.integral)
        df=rbind(df,k)
        writeLines(paste('\n',model,samp,'\n'))
        print('confusion matrix')
        print(res[[2]]$table)
        print('AUC-PR')
        print(res[[3]]$auc.integral)
    }
    print(ggplot(data = df,aes(x=Recall,y=Precision,color=samp))+geom_line()+ggtitle(label = model)+theme_minimal())
    
    return(df %>% mutate(model=model))
}



```
```{r}
data <- read_data('GFP_RPE_CGM.tsv','GO:0007601')
summary(data)
```


I also made a test data set using the BreastCancer data set from mlbench, which is also an unbalanced data set, just to make sure things work the way they should be 

```{r  message=FALSE, results = 'hide'}
data("BreastCancer")
col <- BreastCancer%>%mutate(in_go_set= ifelse(Class=='benign','in_set','not_in_set'),Class=NULL,Id=NULL)%>%pull(in_go_set)
t_data <- BreastCancer%>%mutate(in_go_set= ifelse(Class=='benign','in_set','not_in_set'),Class=NULL,Id=NULL)%>%
    apply(2,as.numeric)%>%as.data.frame%>% mutate(in_go_set=col) #data was a little too
t_data[is.na(t_data)] <- 0

```
```{r }
summary(BreastCancer)
```


## Analysis
the first model I tried was random forests, through the ranger package

test data

```{r}
test <- compare_sampling(data = t_data,model = 'ranger')
```

the CGM data 

```{r}
cgm_dat <- compare_sampling(data = data,model = 'ranger')

```

that clearly did not work well lets test out a new model - glm


```{r}
cgm_bglm <- compare_sampling(data = data,model = 'bayesglm')
```

This also did not work, try lets try the other data sets 

RPE RFP sample - Random Forest

```{r}
rfp_data <- read_data(filename ='RFP_RPE_CGM.tsv', go_id ='GO:0007601' )
rfp_ranger <- compare_sampling(data = rfp_data, model = 'ranger')
```

RPE RFP sample - glm

```{r}
rfp_bglm <- compare_sampling(data = rfp_data, model = 'bayesglm')
```

iPSC sample - Random Forest

```{r}
ipsc_data <- read_data(filename ='IPSC_CGM.tsv', go_id ='GO:0007601' )
ipsc_ranger <- compare_sampling(data = ipsc_data, model = 'ranger')
```

iPSC sample - glm

```{r}
ipsc_bglm <- compare_sampling(data = ipsc_data, model = 'bayesglm')
```








